---
title: "SurvivalAnalysis"
author: "Ryan C. Dallavia"
date: "11/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
data = read.csv("Survival.csv", stringsAsFactors = FALSE)
```

```{r}
str(data)
```
Create right-censoring var (still working at end of study or left org for some reason other than voluntary turn over)
```{r}
data = data %>%
    mutate(censored = if_else((Turnover == 1), 1, 0))
```

```{r}
head(data, n=20)
```
```{r}
hist(data$LOS)
```

```{r}
hist(subset(data, Turnover==0)$LOS)
```

```{r}
# 1 is voluntary turnover
hist(subset(data, Turnover==1)$LOS)
```

#Kaplan-Meir-Analysis (KMA)
```{r}
#install.packages("survival")
library(survival)
```

```{r}
# specify KMA model; empty model implies no covariates; categorical covariates can be included in this type of model. ~1 says gimme null model

# LOS = Time to event
# Censored is second var 
km_fit1 = survfit(Surv(LOS, censored) ~ 1, 
                  data = data,
                  type = "kaplan-meier")
```

```{r}
print(km_fit1)
```
# summarize KM analysis results using default time intervals and create a life table
```{r}
# time (days) the event happens
# n.risk (num of people at risk of event)
# n.event (num of people who actually experience the event - here voluntary turnover)
# survival (% of intervals who passed thru, survived, the event - cumulative survival rate)
summary(km_fit1)
```

# summarize km analysis results by pre-specified time interval & create life table
```{r}
summary(km_fit1, times=c(30, 60, 90*(1:30)))
```

```{r}
# plot cum survival rates
plot(km_fit1)
```

```{r}
install.packages("survminer")
library(survminer)
```

```{r}
# from serveminer package
ggsurvplot(km_fit1, data=data, risk.table = TRUE, conf.int = TRUE,
           ggtheme=theme_classic())

```
# adds covariate (i.e., categorical predictor variable) race; note median equals time to event
```{r}
km_fit2 = survfit(Surv(LOS, censored) ~ Race, 
                  data = data,
                  type = "kaplan-meier")
print(km_fit2)
# n is sample size for each group, events is num of people who experienced event by the end of the study, median time to event, CI intervals around those events.
```

```{r}
summary(km_fit2)
```

```{r}
ggsurvplot(km_fit2, data=data, risk.table = TRUE, conf.int = TRUE,
           ggtheme=theme_classic(), pval=TRUE, pval.method=TRUE)
# pval tells you if curves differ by race, pval.method = TRUE applies log-rank test (prints p-val to see if observed diffs are stat significantly)
```

Cox Prop Hazard Model (cox regressions)
```{r}
cox_reg1 = coxph(Surv(LOS, censored) ~ Race, data=data)
summary(cox_reg1) 
#coef -> euler's num ^coef 
#z=waldstat
#everything interpreted relative to the black group b/c its done alphabetically
#concordance -> how well did the model predicted individ. people's experience of event in terms of time; indicates how well model predicted event. Here, it's no better than a coin toss. SE variable of .014 says including our covariate, race improved our model very little. 
# Log rank test is significance you report with smaller sample sizes.
```
```{r}
cox_reg2 = coxph(Surv(LOS, censored) ~ Race + Pay_sat + Pay_hourly, data=data)
summary(cox_reg2)
```
Second table has hazard coefficients. Hourly coefficient is significant. Below, tells us risk of individual experiencing (voluntarily tourning over) the focal event will decrease by about 10.39% for every $ earned. Conf interval above relates to hazard ratio (.8961) not its reciprocal (1.1160)
```{r}
x = 1 - 0.8961
x
```

All are relative to the black reference group. Below equation simply an expression of the above coefficients.
```{r}
log_overall_risk = .121*race_hsp - .044*race_white -.110*pay_hourly + .141*pay_satisfaction
```

